### 2016-06-26 abstract.tex
edits to secs. 1 - 4

  Serializability is a well-understood mechanism that simplifies
  reasoning about the behavior of concurrent transactions by ensuring
  an executing transaction is \emph{isolated} from the effects of
  other concurrenty executing transactions once it begins executing.
  However, enforcing serializability comes at the cost in performance
  because it necessarily restricts opportunities to exploit
  concurrency even when such opportunities would not violate
  application-specific invariants.

>>>>

  Serializability is a well-understood mechanism that simplifies
  reasoning about the behavior of concurrent transactions by ensuring
  transactions are \emph{isolated} from each other's effects once they
  begin executing.  However, enforcing serializability comes at a
  steep cost in performance because it necessarily restricts
  opportunities to exploit concurrency even when such opportunities
  would not violate application-specific invariants.

### 2016-06-26 ansi-isolation.tex
edits to secs. 1 - 4

\begin{figure*}[t]

@@

Fig.~\ref{fig:ansi-isolation} shows the specification of standard
isolation guarantees expressed as constraints over trace
well-formedness. For brevity and convenience, we adopt few notations
in this and subsequent sections.  An execution trace is destructed
into $\A$ and $\visZ$ whenever individual components of the pair are
needed. Otherwise, it is written as $\E$. Sometimes, the dot notation
(\eg~$\E.\A$) is also used. Since $\A$ and $\visZ$ are both sets, we
lift the operations on sets to pairs of sets when updating $\E$. For
example, $\E' = \E \cup (\{\eta_2\},\{(\eta_1,\eta_2)\})$ expands to
$\E' = (\E.\A \cup \{\eta_2\},\,\E.\visZ \cup \{(\eta_1,\eta_2)\})$.
When $\psi$ is a proposition, $\underE{\psi}$ denotes the
interpretation of $\psi$ in the context of the trace $\E$. Such
interpretations are defined on a case-by-case basis in
Fig.~\ref{fig:ansi-isolation}. In the following, we give informal
explanations for each definition.

In the context of a trace $(\A,\visZ)$, an effect $\eta$ is said to
belong to a transaction $T_i$ if $\eta$ belongs to the effect set $A$
and its transaction identifier is $T_i$. The containment relation is
trivially lifted to the set of effects to define $\underE{S \subseteq
T_i}$.  The interpretation of $\eta_1 \visar \eta_2$ and $\eta_1
\soar \eta_2$ are straightforward and explained in \S\ref{sec:syntax}.
A transaction $T_i$ is said to be visible to an effect $\eta$ if every
effect $\eta_1$ of $T_i$ recorded by the trace is visible to $\eta$.
$T_i$ may be visible to $\eta$ but may not be visible to every other
effect in the transaction. For a transaction $T_i$ to be considered to
be visible to a transaction $T_j$ in the context of a trace $\E$
(written $\underE{T_i \visar T_j}$), every effect ($\eta_1$) of $T_i$
present in $\E$ must be visible to every effect ($\eta_2$) of $T_j$ in
$\E$. Conversely, if none of the effects of $T_i$ present in $\E$ are
visible to any effect of $T_j$, then $T_i$ is considered invisible to
$T_j$ under $\E$ (written $\underE{T_i \invisar T_j}$). Transaction
$T_i$ is said to have written to a variable $X$ under $\E$ (i.e.,
$\underE{T_i \wrstoar X}$) if there exists a write-to-$X$ effect from
$T_i$ in $\E$.

@@

The ANSI SQL 92 standard requires \iso{Read Committted} isolation to
avoid the dirty reads phenomenon, which is achieved by enforcing
$\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$ guarantees. The {\sc rc}
specification is therefore a combination of these two guarantees. The
specification also agrees with the description and
implementation~\cite{bailishat,pldi15} of {\sc rc} for highly
available replicated stores. On relational databases, however, {\sc
rc} has also come to be associated with the $\mathtt{MonotonicVis}$
guarantee.  Nonetheless, $\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$
are sufficient to reason about {\sc rc} isolation on relational stores
too. The combination of these guarantees with the {\sc sc} property of
relational stores (formalized in \S\ref{sec:store-consistency})
automatically leads to the monotonicity guarantee, which probably
explains why {\sc rc} comes with $\mathtt{MonotonicVis}$ on such
stores. On weakly consistent stores however, $\mathtt{AtomicVis}$ and
$\mathtt{CommitVis}$ do not imply $\mathtt{MonotonicVis}$. A stronger
isolation level called \iso{Monotonic Atomic View} ({\sc mav} of
Fig.~\ref{fig:ansi-isolation})~\cite{bailishat,pldi15} was proposed to
explicitly extend {\sc rc} with monotonicity on such stores. Sample
{\sc rc} executions shown in Fig.~\ref{fig:rc-executions} are also
{\sc mav}. \iso{Repeatable Read} spec ({\sc rr}) extends {\sc mav}
with the snapshot visibility guarantee as described before.
\iso{Snapshot Isolation} spec ({\sc si}) extends {\sc rr} with a
one-way serializability guarantee w.r.t the transactions that perform
conflicting writes (i.e., writes to the same shared variable).
Finally, \iso{Serializable} isolation extends one-way
$\mathtt{SnapshotVis}$ to all transactions, including those that do
not conflict\footnote{Fig.~\ref{fig:ansi-isolation} presents slightly
weaker versions of {\sc si} and {\sc ser} specs in the interest of
clarity.}. \GK{ToDo: Figures. Need to figure out a convention to
uniformly represent various isolation levels. Current idea: concurrent
timelines with arrows in both directions.}

In contrast to recent proposals (\emph{e.g.},
~\cite{gotsmanconcur15}), our specifications for {\sc si} and {\sc
  ser} do not necessarily impose a total order among transactions
(conflicting or otherwise). In reality, a total order under {\sc ser}
(resp. {\sc si}) is guaranteed only if the store executes all
transactions under {\sc ser} (resp. {\sc SI}) isolation. Our
specifications admit this possibility, and derive a total order under
the assumption of homogenity. However, databases almost always allow
isolation levels to be configured on a per-transaction basis, allowing
transactions at various isolation levels to coexist.  Specifications
that assume homogenity are incorrect under this setting.

>>>>

\begin{figure*}[t!]

@@

Fig.~\ref{fig:ansi-isolation} shows the specification of standard
isolation guarantees expressed as constraints over trace
well-formedness. For brevity and convenience, we adopt few notations
in this and subsequent sections.  An execution trace is destructed
into $\A$ and $\visZ$ whenever individual components of the pair are
needed. Otherwise, it is written as $\E$. Sometimes, the dot notation
(\eg~$\E.\A$) is also used. Since $\A$ and $\visZ$ are both sets, we
lift the operations on sets to pairs of sets when updating $\E$. For
example, $\E' = \E \cup (\{\eta_2\},\{(\eta_1,\eta_2)\})$ expands to
$\E' = (\E.\A \cup \{\eta_2\},\,\E.\visZ \cup \{(\eta_1,\eta_2)\})$.
When $\psi$ is a formula, $\underE{\psi}$ denotes the
interpretation of $\psi$ in the context of the trace $\E$. Such
interpretations are defined on a case-by-case basis in
Fig.~\ref{fig:ansi-isolation}. In the following, we give informal
explanations for each definition.

In the context of a trace $(\A,\visZ)$, an effect $\eta$ is said to
belong to a transaction $T_i$ if $\eta$ belongs to the effect set $A$
and its transaction identifier is $T_i$. The containment relation is
trivially lifted to the set of effects to define $\underE{S \subseteq
  T_i}$.  The interpretation of $\eta_1 \visar \eta_2$ and $\eta_1
\soar \eta_2$ are as explained earlier.  A transaction $T_i$ is said
to be visible to an effect $\eta$ if every effect $\eta_1$ of $T_i$
recorded by the trace is visible to $\eta$.  $T_i$ may be visible to
$\eta$ but may not be visible to every other effect in the
transaction. For a transaction $T_i$ to be considered to be visible to
a transaction $T_j$ in the context of a trace $\E$ (written
$\underE{T_i \visar T_j}$), every effect ($\eta_1$) of $T_i$ present
in $\E$ must be visible to every effect ($\eta_2$) of $T_j$ in
$\E$. Conversely, if none of the effects of $T_i$ present in $\E$ are
visible to any effect of $T_j$, then $T_i$ is considered invisible to
$T_j$ under $\E$ (written $\underE{T_i \invisar T_j}$). Transaction
$T_i$ is said to have written to a variable $X$ under $\E$ (i.e.,
$\underE{T_i \wrstoar X}$) if there exists a write-to-$X$ effect from
$T_i$ in $\E$.

@@

The ANSI SQL 92 standard requires \iso{Read Committted} isolation to
avoid the dirty reads phenomenon, which is achieved by enforcing
$\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$ guarantees. The {\sc rc}
specification is therefore a combination of these two guarantees. The
specification also agrees with the description and
implementation~\cite{bailishat,pldi15} of {\sc rc} for highly
available replicated stores. On relational databases, however, {\sc
rc} has also come to be associated with the $\mathtt{MonotonicVis}$
guarantee.  Nonetheless, $\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$
are sufficient to reason about {\sc rc} isolation on relational stores
too. The combination of these guarantees with the {\sc sc} property of
relational stores (formalized in \S\ref{sec:store-consistency})
automatically leads to the monotonicity guarantee, which 
explains why {\sc rc} comes with $\mathtt{MonotonicVis}$ on such
stores. On weakly consistent stores however, $\mathtt{AtomicVis}$ and
$\mathtt{CommitVis}$ do not imply $\mathtt{MonotonicVis}$. A stronger
isolation level called \iso{Monotonic Atomic View} ({\sc mav} of
Fig.~\ref{fig:ansi-isolation})~\cite{bailishat,pldi15} has been proposed to
explicitly extend {\sc rc} with monotonicity on such stores. Sample
{\sc rc} executions shown in Fig.~\ref{fig:rc-executions} are also
{\sc mav}. \iso{Repeatable Read} ({\sc rr}) isolation  extends {\sc mav}
with the snapshot visibility guarantee as described before.
\iso{Snapshot Isolation} spec ({\sc si}) extends {\sc rr} with a
one-way serializability guarantee w.r.t the transactions that perform
conflicting writes (i.e., writes to the same shared variable).
Finally, \iso{Serializable} isolation extends one-way
$\mathtt{SnapshotVis}$ to all transactions, including those that do
not conflict\footnote{Fig.~\ref{fig:ansi-isolation} presents slightly
weaker versions of {\sc si} and {\sc ser} specs in the interest of
clarity.}. \GK{ToDo: Figures. Need to figure out a convention to
uniformly represent various isolation levels. Current idea: concurrent
timelines with arrows in both directions.}

%% SJ: Not sure this paragraph is necessary here.
%% In contrast to recent proposals (\emph{e.g.},
%% ~\cite{gotsmanconcur15}), our specifications for {\sc si} and {\sc
%%   ser} do not necessarily impose a total order among transactions
%% (conflicting or otherwise). In reality, a total order under {\sc ser}
%% (resp. {\sc si}) is guaranteed only if the store executes all
%% transactions under {\sc ser} (resp. {\sc si}) isolation. Our
%% specifications admit this possibility, and derive a total order under
%% the assumption of homogenity. However, databases almost always allow
%% isolation levels to be configured on a per-transaction basis, allowing
%% transactions at various isolation levels to coexist.  Specifications
%% that assume homogenity are incorrect under this setting.

### 2016-06-26 introduction.tex
edits to secs. 1 - 4

The tension between serializable transactions, which are easy to
reason about but difficult to implement, and more pragmatic goals but
weaker variants that are driven by performance and availability
considerations, has motivated the development of weaker forms of
transaction isolation, beginning as early as 1976~\cite{gray1976}. The
ANSI SQL 92 standard defines three such weak isolation levels which
are now implemented in many relational and NoSQL databases. Not
surprisingly, weakly-isolated transactions have been found to
significantly outperform serializable transactions on benchmark
suites, both on single-node databases and multi-node replicated
stores~\cite{dbtuningbook,bailishat,bailisvldb}, leading to their
overwhelming adoption. A 2013 study~\cite{bailishotos} of 18 popular
ACID and ``NewSQL'' databases found that only three of them offer
serializability by default, and half, including Oracle 11g, do not
offer it at all.  A 2015 study~\cite{bailisferal} of a large corpus of
database applications finds no evidence that applications manifestly
change the default isolation level offered by the database. Taken
together, these studies make clear that weakly-isolated transactions
are quite prevalent in practice, and strongly-isolated (serializable)
transactions are often eschewed.

Unfortunately, weak isolation admits behaviors that are difficult to
comprehend~\cite{berenson}. To quantify weak isolation anomalies,
Fekete \emph{et al.}~\cite{feketevldb09} have devised and experimented
with a microbenchmark suite that executes transactions under a weakly
isolated \emph{read committed} isolation level - the default level for
8 of the 18 databases studied in~\cite{bailishotos}, and found that 25
out of every 1000 rows in the database violate at least one integrity
constraint. Bailis \emph{et al.}~\cite{bailisferal} rely on Rails'
\emph{uniqueness validation} to maintain uniqueness of records while
serving Linkbench's~\cite{linkbench} insertion workload (6400 records
distributed over 1000 keys; 64 concurrent clients), and end up with
more than 10 duplicate records. Rails relies on database transactions
to validate uniqueness during insertions, which is sensible if
transactions are serializable, but incorrect under the \emph{read
  committed} isolation level used in the experiments. The same study
has found that 13\% of all invariants among 67 open source
Ruby-on-Rails applications are at risk of being violated due to weak
isolation. Indeed, incidents of safety violations due to executing
applications in a weakly-isolated environment have been reported on
web services in production~\cite{starbucksbug, scimedbug}, including
in safety-critical applications such as bitcoin
exchanges~\cite{poloniexbug, bitcoinbug}. While enforcing
serializability for all transactions would be sufficient to avoid
these errors and anomalies, it would likely be an overly conservative
strategy; indeed, 75\% of the invariants studied in~\cite{bailisferal}
were shown to be preserved under some form of weak isolation.  When to
use weak isolation, and in what form, is therefore a prominent
question facing all database programmers.

A major problem with weak isolation is that its semantics in the
context of user programs is not well-understood. The original
proposal~\cite{gray1976} defines multiple ``degrees'' of weak
isolation in terms of implementation details such as the nature and
duration of locks held in each case. The ANSI SQL 92 standard defines
four levels of isolation (including serializability) in terms of
various undesirable \emph{phenomena} (\eg \emph{dirty reads}) each is
required to prevent. While this is an improvement, it requires
programmers to be prescient about the possible ways various
undesirable phenomena might manifest in their applications, and
determine, in each case, if the phenomenon can be allowed without
violating application invariants. This is understandably
hard. Adya~\cite{adyaphd} presents the first formal definitions of
some of well-known isolation levels in the context of a sequentially
consistent (SC) database.  However, there has been little progress
relating Adya's system model to a formal operational semantics or a
proof system that can facilitate rigorous correctness arguments.
Consequently, reasoning about weak isolation remains an error prone
endeavor, with major database vendors continuing to document their
isolation levels primarily in terms of the undesirable
phenomena~\cite{postgresiso, mysqliso, oracleiso} a particular
isolation level may induce; it thus squarely becomes the programmer's
responsibility to determine if their application can yield an
execution which can trigger such phenomena.

Recent results on reasoning about application invariants in the
presence of weak \emph{consistency}~\cite{burckhardt14, redblueosdi,
  redblueatc, ecinec, gotsmanpopl16} address broadly related concerns.
At a high level, reasoning about weak consistency involves reasoning
about the concurrent execution of strongly atomic and isolated
operations against different but consistent states of the database or
data store, where each operation locally preserves application
invariants. In contrast, reasoning about weak isolation involves
reasoning about the concurrent execution of transactions, whose
operations witness different states which, although consistent
individually, may not be obviously reconciled into a consistent global
view.  Moreover, operations may violate invariants as long as the
transaction as a whole preserves them.  Intuitively, the latter
problem reduces to the former if all transactions contain a single
operation. Furthermore, weak isolation can exist independent of weak
consistency, as evident from the presence of weakly isolated
transactions on sequentially consistent relational databases; such
systems existed long before the advent of weakly consistent NoSQL
stores.  Thus, frameworks for reasoning about weak isolation will
necessarily have to generalize the reasoning frameworks developed for
weak consistency in new and important ways.

@@

In this paper, we propose a program logic for weakly-isolated
transactions that realizes this goal.  In particular, we develop a set
of syntax-directed compositional proof rules that allow developers to
build correctness proofs for transactional programs in the presence of
a weakly isolated concurrency control mechanism A key novelty of our
approach is that it is parametric over the \emph{isolation semantics}
of transactions in the program, as well as the \emph{consistency
  semantics} of the underlying store. In concrete terms, this means
that, unlike recent work focussed on formalizing the semantics of weak
consistency~\cite{gotsmanpopl16, redblueatc, ecinec}, our system model
does not assume a minimum or predefined set of consistency or
isolation levels. Instead, our operational semantics admits declarative
specifications of transaction isolation and store consistency levels
(\eg~\cite{pldi15,gotsmanconcur15}), and generates executions that are
guaranteed to satisfy these specifications. The result is a flexible
system model that is general enough to incorporate the semantics of a
range of isolation levels on a variety of stores (\eg the sequentially
consistent store of~\cite{adyaphd}, or the causally consistent store
of~\cite{gotsmanpopl16}, etc).  Our key technical contribution is thus
a set of proof rules that demonstrate that a program with a given
selection of isolation levels for its transactions preserves its 
invariants when executed on a store satisfying certain consistency
guarantees (e.g., causal visibility).

@@

The remainder of the paper is organized as follows.  The next section
provides motivation and background on serializable and weakly-isolated
transactions. \S\ref{sec:opsem} presents an operational semantics for
a core language that supports weakly-isolated transactions, and which
is parameterized over different isolation and store-consistency
notions. \S\ref{sec:proofrules} formalizes the proof systemt that we
use to reason about program invariants, and establishes the soundness
of these rules with respect to the semantics.  We describe the impact
of our reasoning framework in the context of several real-world case
studies in \S\ref{sec:casestudies}.  Related work and conclusions are
given in \S\ref{sec:relatedwork}.

>>>>

The tension between serializable transactions, which are easy to
reason about but difficult to implement, and more pragmatic variants
that are driven by performance and availability considerations, has
motivated the development of weaker forms of transaction isolation,
beginning as early as 1976~\cite{gray1976}. The ANSI SQL 92 standard
defines three such weak isolation levels which are now implemented in
many relational and NoSQL databases. Not surprisingly, weakly-isolated
transactions have been found to significantly outperform serializable
transactions on benchmark suites, both on single-node databases and
multi-node replicated stores~\cite{dbtuningbook,bailishat,bailisvldb},
leading to their overwhelming adoption. A 2013
study~\cite{bailishotos} of 18 popular ACID and ``NewSQL'' databases
found that only three of them offer serializability by default, and
half, including Oracle 11g, do not offer it at all.  A 2015
study~\cite{bailisferal} of a large corpus of database applications
finds no evidence that applications manifestly change the default
isolation level offered by the database. Taken together, these studies
make clear that weakly-isolated transactions are quite prevalent in
practice, and strongly-isolated (serializable) transactions are often
eschewed altogether.

Unfortunately, weak isolation admits behaviors that are difficult to
comprehend~\cite{berenson}. To quantify weak isolation anomalies,
Fekete \emph{et al.}~\cite{feketevldb09} have devised and experimented
with a microbenchmark suite that executes transactions under a
weakly-isolated \emph{read committed} isolation level - the default
level for 8 of the 18 databases studied in~\cite{bailishotos}, and
found that 25 out of every 1000 rows in the database violate at least
one integrity constraint. Bailis \emph{et al.}~\cite{bailisferal} rely
on Rails' \emph{uniqueness validation} to maintain uniqueness of
records while serving Linkbench's~\cite{linkbench} insertion workload
(6400 records distributed over 1000 keys; 64 concurrent clients), and
report discovering more than 10 duplicate records.  Rails relies on
database transactions to validate uniqueness during insertions, which
is sensible if transactions are serializable, but incorrect under the
\emph{read committed} isolation level used in the experiments. The
same study has found that 13\% of all invariants among 67 open source
Ruby-on-Rails applications are at risk of being violated due to weak
isolation. Indeed, incidents of safety violations due to executing
applications in a weakly-isolated environment have been reported on
web services in production~\cite{starbucksbug, scimedbug}, including
in safety-critical applications such as bitcoin
exchanges~\cite{poloniexbug, bitcoinbug}. While enforcing
serializability for all transactions would be sufficient to avoid
these errors and anomalies, it would likely be an overly conservative
strategy; indeed, 75\% of the invariants studied in~\cite{bailisferal}
were shown to be preserved under some form of weak isolation.  When to
use weak isolation, and in what form, is therefore a prominent
question facing all database programmers.

A major problem with weak isolation is that its semantics in the
context of user programs is not well-understood. The original
proposal~\cite{gray1976} defines multiple ``degrees'' of weak
isolation in terms of implementation details such as the nature and
duration of locks held in each case. The ANSI SQL 92 standard defines
four levels of isolation (including serializability) in terms of
various undesirable \emph{phenomena} (\eg \emph{dirty reads}) each is
required to prevent. While this is an improvement, it requires
programmers to be prescient about the possible ways various
undesirable phenomena might manifest in their applications, and
determine, in each case, if the phenomenon can be allowed without
violating application invariants. This is understandably hard,
especially in the absence of any formal underpinning to define weak
isolation semantics.  Adya~\cite{adyaphd} presents the first formal
definitions of some of well-known isolation levels in the context of a
sequentially consistent (SC) database.  However, there has been little
progress relating Adya's system model to a formal operational
semantics or a proof system that can facilitate rigorous correctness
arguments.  Consequently, reasoning about weak isolation remains an
error prone endeavor, with major database vendors continuing to
document their isolation levels primarily in terms of the undesirable
phenomena~\cite{postgresiso, mysqliso, oracleiso} a particular
isolation level may induce, placing the burden squarely on the
programmer to determine application correctness.

Recent results on reasoning about application invariants in the
presence of weak \emph{consistency}~\cite{burckhardt14, redblueosdi,
  redblueatc, ecinec, gotsmanpopl16} address broadly related concerns.
At a high level, reasoning about weak consistency involves reasoning
about the concurrent execution of strongly atomic and fully isolated
operations against different (potentially replicated), but consistent
states of the database or data store, where each operation locally
preserves application invariants. In contrast, reasoning about weak
isolation involves reasoning about the concurrent execution of
transactions, whose operations witness different states which,
although consistent individually, may not be obviously reconciled into
a consistent global view.  Intuitively, the latter problem reduces to
the former if all transactions contain a single
operation. Furthermore, weak isolation can exist independent of weak
consistency, as evident from the presence of weakly isolated
transactions on sequentially consistent relational databases; such
systems existed long before the advent of weakly consistent NoSQL
stores.  Thus, frameworks for reasoning about weak isolation will
necessarily have to generalize the reasoning frameworks developed for
weak consistency in new and important ways.

@@

In this paper, we propose a program logic for weakly-isolated
transactions that realizes this goal.  In particular, we develop a set
of syntax-directed compositional proof rules that allow developers to
build correctness proofs for transactional programs in the presence of
a weakly isolated concurrency control mechanism.  A key novelty of our
approach is that it is parametric over the \emph{isolation semantics}
of transactions in the program, as well as the \emph{consistency
  semantics} of the underlying store. In concrete terms, this means
that, unlike recent work focussed on formalizing the semantics of weak
consistency~\cite{gotsmanpopl16, redblueatc, ecinec}, our system model
does not assume a minimum or predefined set of consistency or
isolation levels. Instead, our operational semantics admits
declarative specifications of transaction isolation and store
consistency levels (\eg~\cite{pldi15,gotsmanconcur15}), and generates
executions that are guaranteed to satisfy these specifications. The
result is a flexible system model that is general enough to
incorporate the semantics of a range of isolation levels on a variety
of stores (\eg the sequentially consistent store of~\cite{adyaphd}, or
the causally consistent store of~\cite{gotsmanpopl16}, etc).  Our key
technical contribution is thus a set of proof rules that demonstrate
that a program with a given selection of isolation levels for its
transactions preserves its invariants when executed on a store
equipped with certain consistency guarantees (e.g., causal
visibility).

@@

The remainder of the paper is organized as follows.  The next section
provides motivation and background on serializable and weakly-isolated
transactions. \S\ref{sec:opsem} presents an operational semantics for
a core language that supports weakly-isolated transactions,
parameterized over different isolation notions. \S\ref{sec:proofrules}
formalizes the proof systemt that we use to reason about program
invariants, and establishes the soundness of these rules with respect
to the semantics. \S\ref{sec:store-consistency} generalizes the
framework to integrate support for weakly consistent data stores.  We
describe the impact of our reasoning framework in the context of
several real-world case studies in \S\ref{sec:casestudies}.  Related
work and conclusions are given in \S\ref{sec:relatedwork}.
